---
title: "Search Feature Experiment and User Engagement Analysis"
output:
  github_document:
    toc: true
    df_print: kable
---

## Overview

This analysis evaluates a synthetic 28-day randomized experiment comparing the existing search experience with an improved-results variant. The assignment unit is the user, so the primary statistical test is performed on user-level success rates rather than treating repeated search sessions as independent observations.

```{r setup, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(knitr)

project_root <- if (file.exists("requirements.txt")) "." else ".."

read_project_csv <- function(file_name) {
  file_path <- file.path(project_root, "data", "raw", file_name)

  if (!file.exists(file_path)) {
    stop(
      paste0(
        "Could not find ", file_path,
        ". Open the repository as an RStudio project or knit from the analysis folder."
      )
    )
  }

  read.csv(
    file_path,
    row.names = NULL,
    check.names = FALSE,
    stringsAsFactors = FALSE
  )
}

assignments <- read_project_csv("experiment_assignments.csv")
sessions <- read_project_csv("search_sessions.csv")
```

## Experiment design

- **Population:** 12,000 synthetic users with at least one search session
- **Randomization unit:** User
- **Control:** Existing search results
- **Treatment:** Improved search results
- **Primary metric:** Mean user search success rate, where a successful session includes a result click followed by at least two minutes of watch time
- **Diagnostics:** Click-through rate, zero-result rate, and watch time after search
- **Guardrail:** Search error rate

Secondary and segment results are exploratory. The primary metric is evaluated with a two-sided Welch t-test at the user level.

## Data checks

```{r data-checks}
data_checks <- data.frame(
  check = c(
    "Unique user assignments",
    "Unique search sessions",
    "Sessions with valid assigned user",
    "Successful sessions with a click",
    "Successful sessions with at least two watch minutes"
  ),
  passed = c(
    !any(duplicated(assignments$user_id)),
    !any(duplicated(sessions$session_id)),
    all(sessions$user_id %in% assignments$user_id),
    all(sessions$clicked_result[sessions$successful_search == 1] == 1),
    all(sessions$watch_minutes_after_search[sessions$successful_search == 1] >= 2)
  )
)

kable(data_checks)
```

## Assignment balance

```{r assignment-balance}
assignment_counts <- table(
  factor(
    assignments$variant,
    levels = c("control", "improved_results")
  )
)

assignment_summary <- data.frame(
  variant = names(assignment_counts),
  assigned_users = as.integer(assignment_counts),
  assignment_share = as.integer(assignment_counts) / sum(assignment_counts)
)

kable(assignment_summary, digits = 3)

srm_test <- chisq.test(assignment_counts, p = c(0.5, 0.5))
srm_test
```

## User-level metrics

```{r user-level-metrics}
user_metrics <- sessions %>%
  group_by(user_id) %>%
  summarise(
    search_sessions = n(),
    success_rate = mean(successful_search),
    click_through_rate = mean(clicked_result),
    zero_result_rate = mean(zero_results),
    error_rate = mean(search_error),
    avg_watch_minutes = mean(watch_minutes_after_search),
    .groups = "drop"
  ) %>%
  inner_join(assignments, by = "user_id")

variant_summary <- user_metrics %>%
  group_by(variant) %>%
  summarise(
    users = n(),
    mean_user_success_rate = mean(success_rate),
    mean_user_click_through_rate = mean(click_through_rate),
    mean_user_zero_result_rate = mean(zero_result_rate),
    mean_user_error_rate = mean(error_rate),
    avg_watch_minutes = mean(avg_watch_minutes),
    .groups = "drop"
  )

kable(variant_summary, digits = 4)
```

## Primary statistical test

```{r primary-test}
control_success <- user_metrics %>%
  filter(variant == "control") %>%
  pull(success_rate)

treatment_success <- user_metrics %>%
  filter(variant == "improved_results") %>%
  pull(success_rate)

primary_test <- t.test(
  treatment_success,
  control_success,
  alternative = "two.sided",
  var.equal = FALSE,
  conf.level = 0.95
)

primary_result <- data.frame(
  control_mean = mean(control_success),
  treatment_mean = mean(treatment_success),
  absolute_lift_percentage_points =
    (mean(treatment_success) - mean(control_success)) * 100,
  relative_lift_pct =
    (mean(treatment_success) / mean(control_success) - 1) * 100,
  ci_lower_percentage_points = primary_test$conf.int[1] * 100,
  ci_upper_percentage_points = primary_test$conf.int[2] * 100,
  p_value = primary_test$p.value
)

kable(primary_result, digits = 4)
primary_test
```

## Guardrail test

```{r guardrail-test}
control_error <- user_metrics %>%
  filter(variant == "control") %>%
  pull(error_rate)

treatment_error <- user_metrics %>%
  filter(variant == "improved_results") %>%
  pull(error_rate)

guardrail_test <- t.test(
  treatment_error,
  control_error,
  alternative = "two.sided",
  var.equal = FALSE,
  conf.level = 0.95
)

guardrail_result <- data.frame(
  control_error_rate = mean(control_error),
  treatment_error_rate = mean(treatment_error),
  difference_percentage_points =
    (mean(treatment_error) - mean(control_error)) * 100,
  ci_lower_percentage_points = guardrail_test$conf.int[1] * 100,
  ci_upper_percentage_points = guardrail_test$conf.int[2] * 100,
  p_value = guardrail_test$p.value
)

kable(guardrail_result, digits = 4)
```

## Platform results

```{r platform-results}
platform_summary <- user_metrics %>%
  group_by(platform, variant) %>%
  summarise(
    users = n(),
    mean_user_success_rate = mean(success_rate),
    mean_user_error_rate = mean(error_rate),
    .groups = "drop"
  )

kable(platform_summary, digits = 4)
```

```{r primary-chart, fig.width=8, fig.height=5}
ggplot(
  variant_summary,
  aes(
    x = factor(
      variant,
      levels = c("control", "improved_results"),
      labels = c("Control", "Improved results")
    ),
    y = mean_user_success_rate,
    fill = variant
  )
) +
  geom_col(width = 0.65, color = "#303134", show.legend = FALSE) +
  geom_text(
    aes(label = scales::percent(mean_user_success_rate, accuracy = 0.1)),
    vjust = -0.5
  ) +
  scale_fill_manual(values = c("#5F6368", "#1A73E8")) +
  scale_y_continuous(
    labels = scales::percent_format(accuracy = 1),
    limits = c(0, 0.60)
  ) +
  labs(
    title = "Mean User Search Success Rate by Experiment Variant",
    subtitle = "Synthetic 28-day randomized experiment; user is the analysis unit",
    x = "Experiment variant",
    y = "Mean user success rate"
  ) +
  theme_minimal()
```

## Interpretation

The experiment should be evaluated using the pre-specified user-level primary metric, assignment-balance check, confidence interval, and search-error guardrail. Platform and query-category results can help identify follow-up questions, but they should not replace the overall randomized comparison or be treated as independently confirmed effects without adjustment for multiple testing.

Because the dataset is synthetic, the result demonstrates an experiment-analysis workflow rather than the performance of a real search product.
